Nonparametric Block Bootstrap Kolmogorov-Smirnov Goodness-of-Fit Test

### 2024-04-07 manuscript/manuscript.tex

\title{\bf Nonparametric Bootstrap Kolmogorov-Smirnov Goodness-of-Fit Test for
  Marginal Distributions of Stationary Time Series}
\if0\blind
{
  \author{Mathew Chandy, %\\
%   \href{mailto:mathew.chandy@uconn.edu}
%   {\nolinkurl{mathew.chandy@uconn.edu}}\\
  Elizabeth Schifano\\
  Jun Yan\\
  Xianyang Zhang\\
} \fi

@@

The Kolmogorov-Smirnov (KS) test is widely used to test if a sample is
from a given distribution. While a non-parametric bootstrap bias correction
exists for independent and identically distributed (i.i.d.) samples with unknown
parameters, a comparable correction suitable 
for serially dependent data is lacking.
This study proposes a bias correction
using non-parametric 
block bootstrap, designed to approximate the KS statistic in scenarios where
parameters are unspecified and data exhibit serial dependence. We provide
theoretical justification for this approach, then we evaluate its size
and power through simulation. Finally, we demonstrate applications of 
the method to
stock returns of the S\&P 100 and S\&P 500 indices.

\bigskip
\noindent{\sc Keywords}:
bias-correction; 
dependent data; 
time series. 
\end{abstract}

@@

Despite its widespread use, the KS test can be 
misapplied when its foundational assumptions are overlooked. The KS test 
assumes the data are independently and identically distributed (i.i.d.) and 
that the hypothesized distribution is continuous and fully specified without 
the need for parameter estimation. \citet{zeimbekakis2022misuses} examines 
common misapplications of the one-sample KS test, notably the inappropriate
use when the hypothesized distribution contains unspecified parameters.
For this situation, \citet{zeimbekakis2022misuses} proposed a parametric
bootstrap, where bootstrap samples of the test statistics are constructed from
samples generated from the fitted hypothesized distribution.
Alternatively, \citet{babu2004goodness} address this issue through the
application of basic non-parametric bootstrap methods.

Another prevalent misapplication of the KS test arises when data exhibit serial dependence, often overlooked in analyses. In an investigation of performance variation in
high-performance computing systems, \citet{tuncer2019ieee} did not describe
how or if they accounted for serial dependence when applying the two-sample
KS test.
\citet{zeimbekakis2022misuses} proposed a semi-parametric bootstrap
where the serial dependence structure is modeled via a working serial copula.

A particularly complex scenario arises when the distribution has unspecified
parameters, and the data are serially dependent. For this,
\citet{zeimbekakis2022misuses} also recommends a semi-parametric bootstrap
approach, where the parameters of the hypothesized distribution and the serial
dependence both need to be estimated and their uncertainty accounted for.

Developing a fully non-parametric solution for the KS 
tests in the presence of serially dependent data presents a significant 
challenge. The null distribution's characteristics are intricately linked to 
the structure of the serial dependence, which can vary widely in practical 
situations. Employing a non-parametric bootstrap would necessitate defining a 
specific model for this dependence, despite the primary focus being on 
evaluating the marginal distribution of a stationary series. To date, a bias 
correction method for block bootstrap, akin to the approach \citet{babu2004goodness} 
applied to the basic bootstrap, has not been established. This study seeks 
to bridge this gap by introducing a bias correction technique for the 
non-parametric block bootstrap, tailored for use in scenarios where the 
hypothesized distribution includes unspecified parameters and the data 
exhibit serial dependence. Our approach reduces to that of
\citet{babu2004goodness} when the data are independent.

@@

\subsection{Review of Kolmogorov-Smirnov Test}

@@

We note that
\begin{equation*}
Y_n(x; \hat\theta) = \sqrt{n}(F_n(x) - F(x)) - 
\sqrt{n}(F(x; \hat\theta_n) - F(x)),
\end{equation*}
where $F(x)$ is the true cdf (under the null $F(x) = F(x, \theta_0)$ for some
true parameter $\theta_0$).

\subsection{Non-Parametric Basic Bootstrap}

@@

The p-value of the basic bootstrap KS test can be computed
as $p = \#\{T^{(b)}_n > T_n\} / B$ for 
$b \in \{1, \ldots, B\}$.

\subsection{Non-Parametric Block Bootstrap}

@@

We then use bootstrap to approximate the distribution of
the KS statistic under $H_0$. In this case, we can 
use the distribution of $\sqrt{n}(F^{(b)}_n(x) - E[F^{(b)}_n(x)])$
as an approximation of the distribution of
$\sqrt{n}(F_n(x) - F(x))$,
and the distribution of 
$\sqrt{n}(F(x; \hat\theta^{(b)}_n) - F(x; E[\hat\theta^{(b)}_n]))$ to
approximate $\sqrt{n}(F(x; \theta_n) - F(x))$'s distribution.

We can compute $E[F^{(b)}_n(x)]$ and
$E[\hat\theta^{(b)}_n]$ numerically (they can also be computed analytically, 
depending on the types of block bootstrap we use). One can compute 
$F^{(b)}_n$ 
and $\hat\theta^{(b)}_n$ based on
the $b$th bootstrap sample. Then
$E[F^{(b)}_n(x)] \approx \frac{1}{B}\sum_{b = 1}^BF^{(b)}_n(x)$, and
$E[\hat\theta^{(b)}_n] \approx \frac{1}{B}\sum_{b = 1}^B\hat\theta^{(b)}_n$.
In this case, we can define
\begin{align*}
  Y^{(b)}_n(x) &= \sqrt{n}(F^{(b)}_n(x) - E[F^{(b)}_n(x)]) - 
             \sqrt{n}(F(x; \hat\theta^{(b)}_n) - F(x; E[\hat\theta^{(b)}_n)]) \\
           &= \sqrt{n}(F^{(b)}_n(x) - F(x; \hat\theta^{(b)}_n)) -
             \sqrt{n}(E[F^{(b)}_n(x)] - F(x; E[\hat\theta^{(b)}_n])),
\end{align*}
and $T^{(b)}_n = \sup_x|Y^{(b)}_n(x)|$. Each $T_n^{(b)}$,
$b =1, \ldots, B$, is considered a draw from a distribution that approximates
the distribution of $T_n$. Therefore, the p-value of the observed statistic
$T_n$ can be assessed by positioning it agains the empirical distribution of
$T_n^{(b)}$, $b = 1, \ldots, B$.

@@

The p-value of the block bootstrap KS test can be computed
as $p = \#\{T^{(b)}_n > T_n\} / B$ for 
$b \in \{1, \ldots, B\}$.

@@

\subsection{Size}
We first must evaluate whether this method works when $H_0$ is true. To
test this, we can
generate a simulated sample $X_t$ with a certain marginal distribution 
$F_\theta$,
and use our method to test if $X_t$ indeed has the marginal distribution $F$ 
with some unknown $\theta$. If the test holds its size, the 
p-value
of the test should be uniformally distributed between 0 and 1. We must try the
method with different marginal distributions to ensure that it is robust.
In order for the method to work, a large sample size may be necessary.

We generated time series with marginal distributions $N(8, 8)$ and
$\Gamma(8, 1)$ with Kendall's
\jy{Need to state how serial dependence were introduced, like you did in the
  first paper: transforming a normal series.}
\mc{addressed}
$\tau \in \{-.75, -.5, -.25, 0, .25, .5, 75\}$, and
sample size $n \in \{100, 200, 400, 800\}$. Kendall's $\tau$ was chosen as a
measure of serial dependence as it does not vary between two different 
marginal distributions.
To generate the samples to which our
method would be applied, we simulated a time series $W_t$ from a 1st 
order autoregressive (AR(1)) process:
\begin{equation*}
W_t = \phi X_{t-1} + \epsilon_t,
\end{equation*}
where $\phi$ is an autoregressive coefficient, and $\epsilon_t$ is a series of
independent errors from a normal distribution with mean zero and variance
$\sigma_{\epsilon}^2$. The strength of the serial dependence is controlled by
$\phi$, which was set to five levels: 
$\{-0.924, -0.707, -0.383, 0, 0.383, 0.707, 0.924\}$, as these
correspond to the desired values for $\tau$. First, we generated a
marginal $N(8, 8)$ by marginally transforming $W_t$ by
\begin{equation*}
X_t = F^{-1}[\Phi(W_t)],
\end{equation*}
where $F^{-1}(p)$ is the quantile function for the $N(8, 8)$ 
distribution.
Then we generated a marginal gamma series by the same procedure, but
replacing $F^{-1}(p)$ with the quantile function for the $\Gamma(8, 1)$
distribution.
After the transformation
to the marginal gamma distribution, the autocorrelations are $\{-0.884, 
-0.686, -0.389, -0.033, 0.327, 0.639, 0.861\}$.

@@

\jy{Consider presenting this in a graph. Try explaining the pattern.
  Keep the notations consistent: $n$ and n are different.}
\mc{addressed}

@@

From 
Figure~\ref{fig:rr}, 
the rejection rates appear to get lower as $\tau$ increases in a
positive direction. Using a sample size $n \geq 800$ appears to be sufficient
when $\tau$ is not too high, but for $\tau \geq 0.75$, all sample sizes 
observed appeared to have less than desirable rejection rates. Results
do not appear to be to different when comparing gamma and normal 
margins, indicating that the method is robust to the marginal distribution.

The rejection rates for $n = 800$
are very 
high (almost 1), indicating that under $H_A$, the test is very powerful 
(low probability of failed rejection $\beta$) when a large enough
sample size is used.

@@

\jy{Quality control the bib source}
\mc{addressed}

>>>>

\title{Nonparametric Block Bootstrap Kolmogorov-Smirnov Goodness-of-Fit Test}
\if0\blind
{
  \author{Mathew Chandy, %\\
%   \href{mailto:mathew.chandy@uconn.edu}
%   {\nolinkurl{mathew.chandy@uconn.edu}}\\
  Elizabeth Schifano\\
  Jun Yan\\
  Xianyang Zhang\\
} \fi

@@

The Kolmogorov-Smirnov (KS) test is a widely used statistical test that
assesses the conformity of a sample to a specified distribution. Its efficacy,
however, diminishes with serially dependent data and when parameters
within the hypothesized distribution are unknown. For independent data,
parametric and nonparametric bootstrap procedure are available to adjust for
estimated parameters. For serially dependent stationary data, parametric
bootstrap has been developed with a working serial dependence structure. A
counterpart for the nonparametric bootstrap approach, which needs a bias
correction, has not been studied. Addressing this gap, our study introduces a
bias correction method employing a non-parametric block bootstrap, which
adeptly the distribution of the KS statistic for by unspecified serial
dependence and unspecified parameters. We assess its effectiveness
through simulations, scrutinizing both its size and power. The practicality of
our method is further illustrated with an examination of stock returns from the
S\&P 100 and S\&P 500 indices, showcasing its utility in real-world
applications.

\bigskip
\noindent{\sc Keywords}:
bias-correction; 
serial dependence;
time series. 
\end{abstract}

@@

Despite its widespread use, the KS test can be 
misapplied when its foundational assumptions are overlooked. The KS test 
assumes the data are independently and identically distributed (i.i.d.) and 
\jy{no defining acronyms that are not used later}
that the hypothesized distribution is continuous and fully specified without 
the need for parameter estimation. \citet{zeimbekakis2022misuses} examines
common misapplications of the one-sample KS test. One notable inappropriate
use is when the hypothesized distribution contains unspecified parameters.
In this case, the tests are generally constructed by substituting the unknown
parameters by their estimates, and the asymptotic null distribution of
the test statistic may depend in a complex way on the unknown parameters.
The problem of unspecified parameters can be handled by a parametric
bootstrap, where bootstrap samples of the test statistics are constructed from
samples generated from the fitted hypothesized distribution.
Alternatively, \citet{babu2004goodness} address this issue through a
non-parametric bootstrap method that corrects the bias in the asymptotic null
distribution.

Another prevalent misapplication of the KS test arises when data exhibit serial
dependence, which is often overlooked in analyses. For example, in an
investigation of performance variation in high-performance computing systems,
\citet{tuncer2019ieee} did not describe how or if they accounted for serial
dependence when applying the two-sample KS test.
\citet{zeimbekakis2022misuses} proposed a semi-parametric bootstrap
where the serial dependence structure is modeled via a working serial copula.
A particularly complex scenario arises when the distribution has unspecified
parameters, and the data are serially dependent. For this,
\citet{zeimbekakis2022misuses} also recommends a semi-parametric bootstrap
approach, where the parameters of the hypothesized distribution and the serial
dependence both need to be estimated and their uncertainty accounted for.

Developing a fully non-parametric solution for the KS 
tests in the presence of serially dependent data presents a significant 
challenge. The null distribution's characteristics are intricately linked to 
the structure of the serial dependence, which can vary widely in practical 
situations. Employing a non-parametric bootstrap would necessitate defining a 
specific model for this dependence, despite the primary focus being on 
evaluating the marginal distribution of a stationary series. To date, a bias 
correction method for block bootstrap, akin to the non-parametric bootstrap
approach of \citet{babu2004goodness} for independent data,
has not been established. This study seeks 
to bridge this gap by introducing a bias correction technique for the 
non-parametric block bootstrap, tailored for use in scenarios where the 
hypothesized distribution includes unspecified parameters and the data 
exhibit serial dependence. Our approach reduces to that of
\citet{babu2004goodness} when the data are independent.

@@

\subsection{Kolmogorov--Smirnov Test}

@@

\jy{State clearly the test statistic definition. Point out that the null
  distribution depends on the unknown parameters and unknown serial
  dependence. Point out the challenges are to approximate the null distribution.}

We note that
\begin{equation*}
Y_n(x; \hat\theta) = \sqrt{n}(F_n(x) - F(x)) - 
\sqrt{n}(F(x; \hat\theta_n) - F(x)),
\end{equation*}
\jy{clean up the notations. Has $F$ been defined yet? do you need cdf?}
where $F(x)$ is the true cdf (under the null $F(x) = F(x, \theta_0)$ for some
true parameter $\theta_0$).

\subsection{Non-Parametric Bootstrap for Independent Data}

@@

The p-value of the basic bootstrap KS test can be approximated
as $p = \sum_{b=1}^BB I\{T^{(b)}_n > T_n\} / B$.

\subsection{Non-Parametric Block Bootstrap for Stationary Series}

@@

Continuing using the notation in the independent case,
we use $F^{(b)}_n$ and $\hat\theta^{(b)}_n$ for the empirical distribution and
the estimated parameters based on the $b$th bootstrap sample,
$b = 1, \ldots, B$.
We then use block bootstrap to approximate the asymptotic distribution of
the KS statistic under $H_0$. In particular, we
use the distribution of $\sqrt{n}(F^{(b)}_n(x) - E[F^{(b)}_n(x)])$
as an approximation of the distribution of
$\sqrt{n}(F_n(x) - F(x))$, and the distribution of 
$\sqrt{n}(F(x; \hat\theta^{(b)}_n) - F(x; E[\hat\theta^{(b)}_n]))$ as
an approximation of the distribution of $\sqrt{n}(F(x; \theta_n) - F(x))$.
The expected values $E[F^{(b)}_n(x)]$ and
$E[\hat\theta^{(b)}_n]$ can be approximated by, respectively,
$E_n[F^{(b)}_n(x)] = \frac{1}{B}\sum_{b = 1}^BF^{(b)}_n(x)$ and
$E_n[\hat\theta^{(b)}_n]  =  \frac{1}{B}\sum_{b = 1}^B\hat\theta^{(b)}_n$.

Then, we can define
\begin{align*}
  Y^{(b)}_n(x) &= \sqrt{n}(F^{(b)}_n(x) - E[F^{(b)}_n(x)]) - 
             \sqrt{n}(F(x; \hat\theta^{(b)}_n) - F(x; E[\hat\theta^{(b)}_n)]) \\
           &= \sqrt{n}(F^{(b)}_n(x) - F(x; \hat\theta^{(b)}_n)) -
             \sqrt{n}(E[F^{(b)}_n(x)] - F(x; E[\hat\theta^{(b)}_n])),
\end{align*}
and $T^{(b)}_n = \sup_x|Y^{(b)}_n(x)|$. Each $T_n^{(b)}$,
$b =1, \ldots, B$, is considered a draw from a distribution that approximates
the distribution of $T_n$. Therefore, the p-value of the observed statistic
$T_n$ can be assessed by positioning it agains the empirical distribution of
$T_n^{(b)}$, $b = 1, \ldots, B$.

@@

The p-value of the block bootstrap KS test can be approximated
as $p = \sum_{b=1}^N I\{T^{(b)}_n > T_n\} / B$.

@@

\subsection{Size}
We first must evaluate whether this method works when $H_0$ is true. To
test this, we can
generate a simulated sample $X_t$ with a certain marginal distribution 
$F_\theta$,
and use our method to test if $X_t$ indeed has the marginal distribution $F$ 
with some unknown $\theta$. If the test holds its size, the 
p-value
of the test should be uniformly distributed between 0 and 1. We must try the
method with different marginal distributions to ensure that it is robust.
In order for the method to work, a large sample size may be necessary.

We generated time series with marginal distributions $N(8, 8)$ and
$\Gamma(8, 1)$ with Kendall's
$\tau \in \{-.75, -.5, -.25, 0, .25, .5, 75\}$, and
sample size $n \in \{100, 200, 400, 800\}$. Kendall's $\tau$ was chosen as a
measure of serial dependence as it does not vary between two different 
marginal distributions.
To generate the samples to which our
method would be applied, we simulated a time series $W_t$ from a 1st 
order autoregressive (AR(1)) process:
\begin{equation*}
W_t = \phi X_{t-1} + \epsilon_t,
\end{equation*}
where $\phi$ is an autoregressive coefficient, and $\epsilon_t$ is a series of
independent errors from a normal distribution with mean zero and variance
$\sigma_{\epsilon}^2$. The strength of the serial dependence is controlled by
$\phi$, which was set to five levels: 
$\{-0.924, -0.707, -0.383, 0, 0.383, 0.707, 0.924\}$, as these
correspond to the desired values for $\tau$. First, we generated a
marginal $N(8, 8)$ by marginally transforming $W_t$ by
\begin{equation*}
X_t = F^{-1}[\Phi(W_t)],
\end{equation*}
\jy{define $\Phi$}
\jy{The language issues are more than I expected. Please do a careful proofread,
  now that you are more experienced.}
where $F^{-1}(p)$ is the quantile function for the $N(8, 8)$ 
distribution.
Then we generated a marginal gamma series by the same procedure, but
replacing $F^{-1}(p)$ with the quantile function for the $\Gamma(8, 1)$
distribution.
After the transformation
to the marginal gamma distribution, the autocorrelations are $\{-0.884, 
-0.686, -0.389, -0.033, 0.327, 0.639, 0.861\}$.

@@

@@

From 
Figure~\ref{fig:rr}, 
the rejection rates appear to get lower as $\tau$ increases in a
positive direction. Using a sample size $n \geq 800$ appears to be sufficient
when $\tau$ is not too high, but for $\tau \geq 0.75$, all sample sizes 
observed appeared to have less than desirable rejection rates. Results
do not appear to be to different when comparing gamma and normal 
margins, indicating that the method is robust to the marginal distribution.
The rejection rates for $n = 800$ are very 
high (almost 1), indicating that under $H_A$, the test is very powerful 
(low probability of failed rejection $\beta$) when a large enough
sample size is used.

@@
